---
title: "GeoChron"
output: html_notebook
---
Assess variability of age or CPE estimates at multiple levels.
At the highest level is a region which is composed of $nL$ localities.

The mean age across a region is $e$ which generates means $c_l$ for each locality.

$$ c_l = e + \zeta_c, l = 1,\ldots, nL, \zeta_c \sim N(0, \sigma_c^2)$$
Each locality has a number $nR_l$ of constituent reefs and a mean $c_l$ which generates means for each reef.

$$r_{kl} = c_l + \zeta_r, k = 1,\ldots, nR_l, l = 1\ldots nL, \zeta_r \sim N(0,\sigma_r^2)$$


Each reef has a number $nH_{kl}$ of constituent holes (typically 3 for our data) and the reef mean generates observations for each hole.

$$ x_{jkl} = r_{kl} + \zeta_h, j = 1,\ldots, nH_{kl}, k = 1,\ldots, nR_l, l = 1,\ldots, nL, \zeta_h \sim N(0,\sigma_h^2)$$
Finally, for each hole there are actuall two depths and one would expect the deeper sample to be older than the sample above it. We model the offset $d$ and report its posterior distribution.
We are interested in comparing its variance $\sigma_d^2$ with the other sigmas. 

$$ y_{jkl} = d + x_{jkl} $$
$$ y_{jkl}-x_{jkl} \sim N(\mu_d, \sigma_d^2)  $$

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(magrittr))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(cmdstanr))
suppressPackageStartupMessages(library(bayesplot))
suppressPackageStartupMessages(library(data.table))
```
 
Create data list from Geochronology "sampsum" data table created from "HOBS-geochronology-ms2.rmd". For a simulation illustrating how this model works, see "GeoChron_sim.Rmd."

```{r}
sampsum  <- fread(here::here('sampleSummary_2022-08-08.csv'))
sampsum[ , locality := str_replace(sample, "_R.H.S.", "")]
sampsum[ , locality := str_replace(locality, "JI-WC", "JI")]
sampsum[, locality := fcase(locality %in% setdiff(unique(sampsum$locality), c("GI", "HC", "LS")), locality,
                            locality == "GI", "GI-EC",
                            locality == "HC", "HC-MC",
                            locality == "LS", "LSG")]
sampsum[, locality := factor(locality, levels = c("LSG", "GI-EC", "LC", "LB", "HC-MC", "NP", "BH", "JI", "PC", "MR", "GR"))]
sampsum$reefID <- paste0(sampsum$locality, " R", sampsum$reef)
sampsum$holeID <- stringr::str_extract(sampsum$sample, ".._R.H.")
sampsum$holeN <- stringr::str_extract(sampsum$holeID, "R.H.")
sampsum$holeN <- stringr::str_extract(sampsum$holeN, "H.")
sampsum$holeN2 <- str_replace(sampsum$holeN, "H", "Sample hole ")
sampsum$holeN3 <- str_replace(sampsum$holeN, "H", "Hole ")
sampsum$median_age_bp <- abs(sampsum$Sample_median_age - 2019)
sampsum[, reef2 := paste0("Reef ", reef)]
sampsum[, Sample_n := factor(Sample_n, levels = c(2, 3, 4, 5, 6, 7))]
sampsum[, reef_lab := fcase(locality == "GI-EC" & reef == 4, 3L,
                            locality == "LC" & reef == 5, 4L,
                            locality == "MR" & reef == 2, 1L,
                            locality == "MR" & reef == 3, 2L)][is.na(reef_lab), reef_lab := reef]

sampsum[, depth := str_extract(depth, "..-..")]
setorder(sampsum, "locality", "reef", "holeN", "depth")

sampsum[, Loc := str_sub(sample, 1, 2), by = sample]
sampsum[, Locality := as.integer(factor(Loc, levels = unique(sampsum$Loc)))]
sampsum[, reef3 := as.integer(factor(reefID, levels = unique(sampsum$reefID)))]
sampsum[, reef2Locality := Locality, by = c("Locality", "reef")]
sampsum[, holeID2 := holeID]
sampsum[, sampsperhole := length(holeID), by = holeID2]


sampsum_r2l <- distinct(sampsum[, .(reef, 
                                    locality, 
                                    reefID,
                                    Locality,
                                    reef3,
                                    reef2Locality)])

sampsum_15to25cm <- distinct(sampsum[depth == "15-25", .(sample,
                                                         Sample_median_age,
                                                         median_age_bp,
                                                         Sample_corrected_posterior_age_estimate,
                                                         reef,
                                                         reef_lab,
                                                         locality,
                                                         reefID,
                                                         holeID,
                                                         holeID2,
                                                         Locality,
                                                         reef3,
                                                         reef2Locality)])

sampsum_25to35cm <- distinct(sampsum[depth == "25-35", .(sample,
                                                         Sample_median_age,
                                                         median_age_bp,
                                                         Sample_corrected_posterior_age_estimate,
                                                         reef,
                                                         reef_lab,
                                                         locality,
                                                         reefID,
                                                         holeID,
                                                         holeID2,
                                                         Locality,
                                                         reef3,
                                                         reef2Locality)])

depthdiffs <- sampsum[sampsperhole == 2, .(sample, 
                                           Sample_median_age,
                                           median_age_bp,
                                           Sample_corrected_posterior_age_estimate, 
                                           depth, 
                                           reef,
                                           reef_lab,
                                           locality, 
                                           reefID, 
                                           holeID,
                                           holeID2,
                                           Locality,
                                           reef3,
                                           reef2Locality)]

depthdiffs[, `:=` (depthdiff_mabp = median_age_bp[depth == "25-35"] -  median_age_bp[depth == "15-25"], 
                   depthdiff_cpe = Sample_corrected_posterior_age_estimate[depth == "25-35"] - Sample_corrected_posterior_age_estimate[depth == "15-25"]), by = holeID2]

depthdiffs2 <- distinct(depthdiffs[, .(reef, locality, reefID, holeID, Locality, reef3, reef2Locality, depthdiff_mabp, depthdiff_cpe)])

locs <- distinct(sampsum_15to25cm[, .(locality, Locality)])
locs[, loclab := 11 - Locality] #Changed from 12 to 11 because LB is temporarily removed due to dead shells being used in the Dead C calculation.

reefs <- distinct(sampsum_15to25cm[, .(locality, reefID, reef, reef_lab, reef3)])
#reefs[, `:=` (locality = factor(locality, levels = c("LSG", "GI-EC", "LC", "LB", "HC-MC", "NP", "BH", "JI", "PC", "MR", "GR")), reeflab = 32 - reef3)]

#median estimated sample ages and corresponding time-averaging estimates from Dominguez et al. (2016)
Detal <- data.table(median_age_bp = c(123, 283, 146, 148, 160, 2043), 
                    Sample_corrected_posterior_age_estimate = c(68, 2670, 16, 27, 16, 1937),
                    Locality = c(1, 1, 1, 1, 1, 1),
                    Reef = c(1, 2, 3, 4, 5, 6),
                    Reef2Locality = c(1, 1, 1, 1, 1, 1),
                    nL = 1,
                    nR = 6)

#for median age over space (15-25cm burial depth); scaled version produces cleaner MCMC diagnostics, but is difficult to compare with Dominguez et al. (2016) results
HOBS_medagexspace_15to25cm <- list(N = nrow(sampsum_15to25cm),
                                   Locality = sampsum_15to25cm$Locality,
                                   Reef = sampsum_15to25cm$reef3,
                                   Reef2Locality = sampsum_r2l$reef2Locality,
                                   x = sampsum_15to25cm$median_age_bp,
                                   nL = length(unique(sampsum_15to25cm$Locality)),
                                   nR = length(unique(sampsum_15to25cm$reef3)))

HOBS_medagexspace_15to25cm_scaled <- list(N = nrow(sampsum_15to25cm),
                                          Locality = sampsum_15to25cm$Locality,
                                          Reef = sampsum_15to25cm$reef3,
                                          Reef2Locality = sampsum_r2l$reef2Locality,
                                          x = as.numeric(scale(sampsum_15to25cm$median_age_bp)),
                                          nL = length(unique(sampsum_15to25cm$Locality)),
                                          nR = length(unique(sampsum_15to25cm$reef3)))

#for CPE over space (15-25cm burial depth); scaled version produces cleaner MCMC diagnostics, but is difficult to compare with Dominguez et al. (2016) results
HOBS_cpexspace_15to25cm <- list(N = nrow(sampsum_15to25cm),
                                Locality = sampsum_15to25cm$Locality,
                                Reef = sampsum_15to25cm$reef3,
                                Reef2Locality = sampsum_r2l$reef2Locality,
                                x = sampsum_15to25cm$Sample_corrected_posterior_age_estimate,
                                nL = length(unique(sampsum_15to25cm$Locality)),
                                nR = length(unique(sampsum_15to25cm$reef3)))

HOBS_cpexspace_15to25cm_scaled <- list(N = nrow(sampsum_15to25cm),
                                       Locality = sampsum_15to25cm$Locality,
                                       Reef = sampsum_15to25cm$reef3,
                                       Reef2Locality = sampsum_r2l$reef2Locality,
                                       x = as.numeric(scale(sampsum_15to25cm$Sample_corrected_posterior_age_estimate)),
                                       nL = length(unique(sampsum_15to25cm$Locality)),
                                       nR = length(unique(sampsum_15to25cm$reef3)))

#for median age over space (25-35cm burial depth); scaled version produces cleaner MCMC diagnostics, but is difficult to compare with Dominguez et al. (2016) results
HOBS_medagexspace_25to35cm <- list(N = nrow(sampsum_25to35cm),
                                   Locality = sampsum_25to35cm$Locality,
                                   Reef = sampsum_25to35cm$reef3,
                                   Reef2Locality = sampsum_r2l$reef2Locality,
                                   x = sampsum_25to35cm$median_age_bp,
                                   nL = length(unique(sampsum_25to35cm$Locality)),
                                   nR = length(unique(sampsum_25to35cm$reef3)))

HOBS_medagexspace_25to35cm_scaled <- list(N = nrow(sampsum_25to35cm),
                                          Locality = sampsum_25to35cm$Locality,
                                          Reef = sampsum_25to35cm$reef3,
                                          Reef2Locality = sampsum_r2l$reef2Locality,
                                          x = as.numeric(scale(sampsum_25to35cm$median_age_bp)),
                                          nL = length(unique(sampsum_25to35cm$Locality)),
                                          nR = length(unique(sampsum_25to35cm$reef3)))

#for CPE over space (25-35cm burial depth)
HOBS_cpexspace_25to35cm <- list(N = nrow(sampsum_25to35cm),
                                Locality = sampsum_25to35cm$Locality,
                                Reef = sampsum_25to35cm$reef3,
                                Reef2Locality = sampsum_r2l$reef2Locality,
                                x = sampsum_25to35cm$Sample_corrected_posterior_age_estimate,
                                nL = length(unique(sampsum_25to35cm$Locality)),
                                nR = length(unique(sampsum_25to35cm$reef3)))

#for median age depth differences over space
HOBS_medageddxspace <- list(N = nrow(depthdiffs2),
                            Locality = depthdiffs2$Locality,
                            Reef = depthdiffs2$reef3,
                            Reef2Locality = sampsum_r2l$reef2Locality,
                            x = depthdiffs2$depthdiff_mabp,
                            nL = length(unique(depthdiffs2$Locality)),
                            nR = length(unique(depthdiffs2$reef3)))

#for CPE depth differences over space
HOBS_cpeddxspace <- list(N = nrow(depthdiffs2),
                         Locality = depthdiffs2$Locality,
                         Reef = depthdiffs2$reef3,
                         Reef2Locality = sampsum_r2l$reef2Locality,
                         x = depthdiffs2$depthdiff_cpe,
                         nL = length(unique(depthdiffs2$Locality)),
                         nR = length(unique(depthdiffs2$reef3)))


```

```{r}

#For HOBS data
model_HOBS <- cmdstan_model("GeoChron.stan")

#For Dominguez et al. (2016) data
model_Detal <- cmdstan_model("GeoChron_Detal.stan")

```

```{r}

datalists <- c(as.name("HOBS_medagexspace_15to25cm"), 
               as.name("HOBS_cpexspace_15to25cm"),
               as.name("HOBS_medagexspace_25to35cm"),
               as.name("HOBS_cpexspace_25to35cm"),
               as.name("HOBS_medageddxspace"),
               as.name("HOBS_cpeddxspace"))

for(Data in datalists){
  if(str_detect(Data, "HOBS")){
    model <- cmdstan_model("GeoChron.stan")
  } else{
    model <- cmdstan_model("GeoChron_Detal.stan")
  }
  
  fit = model$sample(data = eval(Data),
        seed = 1234,
        chains = 8,
        parallel_chains = 8,
        refresh = 1000,
        iter_warmup = 5000, #3000,
        iter_sampling = 15000,
        thin =  5,
        adapt_delta = 0.99,
        max_treedepth = 15)

  
  fit$save_output_files("StanOutput")
  fitdiag <- fit$cmdstan_diagnose()
  fwrite(fitdiag, file = here::here(paste0("StanOutput/", Data, "_diagnostics.txt")))
  dim(fit$draws()) # accessing draws seems to fill the structure with draws
  saveRDS(fit, file = here::here(paste0("StanOutput/", Data, ".rds")))
  
}
  
  
  plot1 <- bayesplot::mcmc_intervals(fit$draws("mu_locality")) 
  plot1 <- plot1 +
    geom_point(data = Data, aes(x = x, y = Locality - 12), color = "red") +
    geom_text(data = locs, aes(x = min(plot1$data$ll) - 3, y = loclab, label = locality), inherit.aes = FALSE) +
    xlim(min(plot1$data$ll) - 5, max(plot1$data$hh))
  
  saveRDS(plot1, here::here(paste0("StanOutput/", Data, "locintplot.rds")))
  ggsave(filename = here::here(paste0("StanOutput/", Data, "locintplot.pdf")),
         plot = plot1,
         width = 6,
         height = 6,
         units = "in",
         dpi = 400)
  
  plot2 <- bayesplot::mcmc_intervals(fit$draws("mu_reef"))
  plot2 <- plot2 +
    geom_text(data = reefs, aes(x = min(plot2$data$ll) - 10, y = reeflab, label = reefID), inherit.aes = FALSE) +
    xlim(min(plot2$data$ll) - 12, max(plot2$data$hh))
  
  saveRDS(plot2, here::here(paste0("StanOutput/", Data, "reefintplot.rds")))
  ggsave(filename = here::here(paste0("StanOutput/", Data, "reefintplot.pdf")),
         plot = plot2,
         width = 6,
         height = 6,
         units = "in",
         dpi = 400)


```


```{r}

bayesplot::mcmc_dens(fit$draws("mu_locality"), facet_args = list(scales = "fixed", ncol = 3))
plot1 <- bayesplot::mcmc_intervals(fit$draws("mu_locality")) 
plot1 +
  geom_text(data = locs, aes(x = min(plot1$data$ll) - 3, y = loclab, label = locality), inherit.aes = FALSE) +
  xlim(min(plot1$data$ll) - 5, max(plot1$data$hh))
  

```

```{r}
plot2 <- bayesplot::mcmc_intervals(fit$draws("mu_reef"))
plot2 +
  geom_text(data = reefs, aes(x = min(plot2$data$ll) - 8, y = reeflab, label = reefID), inherit.aes = FALSE) +
  xlim(min(plot2$data$ll) - 10, max(plot2$data$hh))
```

 
 
 

