---
title: "GeoChron"
output: html_notebook
---
Assess variability of age or CPE estimates at multiple levels.
At the highest level is a region which is composed of $nL$ localities.

The mean age across a region is $e$ which generates means $c_l$ for each locality.

$$ c_l = e + \zeta_c, l = 1,\ldots, nL, \zeta_c \sim N(0, \sigma_c^2)$$
Each locality has a number $nR_l$ of constituent reefs and a mean $c_l$ which generates means for each reef.

$$r_{kl} = c_l + \zeta_r, k = 1,\ldots, nR_l, l = 1\ldots nL, \zeta_r \sim N(0,\sigma_r^2)$$


Each reef has a number $nH_{kl}$ of constituent holes (typically 3 for our data) and the reef mean generates observations for each hole.

$$ x_{jkl} = r_{kl} + \zeta_h, j = 1,\ldots, nH_{kl}, k = 1,\ldots, nR_l, l = 1,\ldots, nL, \zeta_h \sim N(0,\sigma_h^2)$$
Finally, for each hole there are actuall two depths and one would expect the deeper sample to be older than the sample above it. We model the offset $d$ and report its posterior distribution.
We are interested in comparing its variance $\sigma_d^2$ with the other sigmas. 

$$ y_{jkl} = d + x_{jkl} $$
$$ y_{jkl}-x_{jkl} \sim N(\mu_d, \sigma_d^2)  $$

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(magrittr))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(cmdstanr))
```
 
 Let's simulate a data set from this model.
 
 Generate a number of localities from the Region (we are thinking an average of 11)
 We also pick a regional mean age, c
 
```{r}
set.seed(1234)
  nL = 11
  c = 30
```

Generate means for each locality
$\sigma_c = 3$
```{r}
locality_sigma = 5
c_l = c + rnorm(n = nL, mean = 0, s = locality_sigma)
```

Each locality has somewhere between 2 to 4 reefs.

$nR$ tells how many reefs for each locality

```{r}
nR = round(runif(n = nL, min = 2, max = 4))

```

```{r}
n = sum(nR)
nRL = data.frame(Locality = integer(n), Reef = integer(n) )
kk = 1
for( ii in 1:nL){
   for( jj in 1:nR[ii]){
     nRL$Locality[kk] = ii
     nRL$Reef[kk] = kk #jj need a unique identifier for each reef
     kk = kk + 1}
}
rm(n)
```
Add locality and reef means
```{r}
reef_sigma = 2
nRL %<>% mutate(locality_mean = 0, reef_mean = 0)
kk = 1
for( ii in 1:nL){
   for( jj in 1:nR[ii]){
     nRL$locality_mean[kk] = c_l[ii]
     nRL$reef_mean[kk] = rnorm(n = 1, m = c_l[ii], s = reef_sigma )
     kk = kk + 1}
}
```

Each reef has three holes and the values for those holes are generate from the reef mean

```{r}
nRL %>% slice(rep(1:n(), each = 3)) -> nHRL # this triplicates the rows
hole_sigma = 2
depth_mean = 4; depth_sigma = 1
nHRL %<>%  mutate( x = 0, y = 0)

for( ii in 1:nrow(nHRL)){
   nHRL$x[ii] = rnorm(n = 1, m = nHRL$reef_mean[ii], s = hole_sigma)
   d = rnorm(n = 1, m = depth_mean, s = depth_sigma)
   nHRL$y[ii] = d + nHRL$x[ii] 
}

```

For completeness, add the Regional mean, c (wich in fact is the mean of the depth 1 age in a hole)

```{r}
nHRL %<>% mutate(regional_mean = c)
```


Create data list from Geochronology "sampsum" data table created from "HOBS-geochronology-ms2.rmd"

```{r}
sampsum[, Loc := str_sub(sample, 1, 2), by = sample]
sampsum[, Locality := as.integer(factor(Loc, levels = unique(sampsum$Loc)))]
# sampsum[, reef2 := fcase(Locality %in% c(1, 2, 5:7, 9:11), reef,
#                          Locality == 3, ifelse(reef == 5, 4L, reef),
#                          Locality == 4, ifelse(reef == 4, 3L, reef),
#                          Locality == 8, as.integer(reef - 1)), by = c("Locality", "reef")]
sampsum[, reef3 := as.integer(factor(reefID, levels = unique(sampsum$reefID)))]
sampsum[, reef2Locality := Locality, by = c("Locality", "reef")]
sampsum[, holeID2 := holeID]
sampsum[, sampsperhole := length(holeID), by = holeID2]

depthdiffs <- sampsum[sampsperhole == 2, .(sample, 
                                           Sample_median_age,
                                           median_age_bp,
                                           Sample_corrected_posterior_age_estimate, 
                                           depth, 
                                           reef, 
                                           locality, 
                                           reefID, 
                                           holeID,
                                           holeID2,
                                           Locality,
                                           reef3,
                                           reef2Locality)]

depthdiffs[, `:=` (depthdiff_mabp = median_age_bp[depth == "25-35"] -  median_age_bp[depth == "15-25"], 
                   depthdiff_cpe = Sample_corrected_posterior_age_estimate[depth == "25-35"] - Sample_corrected_posterior_age_estimate[depth == "15-25"]), by = holeID2]

depthdiffs2 <- distinct(depthdiffs[, .(reef, locality, reefID, holeID, Locality, reef3, reef2Locality, depthdiff_mabp, depthdiff_cpe)])


xtabs(~Locality + Loc, sampsum)

#for median age over space
Data <- list(N = nrow(sampsum),
             Locality = sampsum$Locality,
             Reef = sampsum$reef3,
             Reef2Locality = sampsum$reef2Locality,
             x = sampsum$median_age_bp,
             nL = length(unique(sampsum$Locality)),
             nR = length(unique(sampsum$reef3)))

#for CPE over space
Data <- list(N = nrow(sampsum),
             Locality = sampsum$Locality,
             Reef = sampsum$reef3,
             Reef2Locality = sampsum$reef2Locality,
             x = sampsum$Sample_corrected_posterior_age_estimate,
             nL = length(unique(sampsum$Locality)),
             nR = length(unique(sampsum$reef3)))

#for median age depth differences over space
Data <- list(N = nrow(depthdiffs2),
             Locality = depthdiffs2$Locality,
             Reef = depthdiffs2$reef3,
             Reef2Locality = depthdiffs2$reef2Locality,
             x = depthdiffs2$depthdiff_mabp,
             nL = length(unique(depthdiffs2$Locality)),
             nR = length(unique(depthdiffs2$reef3)))

#for CPE depth differences over space
Data <- list(N = nrow(depthdiffs2),
             Locality = depthdiffs2$Locality,
             Reef = depthdiffs2$reef3,
             Reef2Locality = depthdiffs2$reef2Locality,
             x = depthdiffs2$depthdiff_cpe,
             nL = length(unique(depthdiffs2$Locality)),
             nR = length(unique(depthdiffs2$reef3)))

```



Now let's see if Stan can recover these values

```{r}

Data = list( N = nrow(nHRL),
             Locality = nHRL$Locality,
             Reef = nHRL$Reef,
             Reef2Locality = nRL$Locality, 
             x = nHRL$x,
             y = nHRL$y,
             nL = length(unique(nHRL$Locality)),
             nR = length(unique(nHRL$Reef)))

model <- cmdstan_model("GeoChron.stan")

```

```{r}

fit = model$sample(data = Data,
    seed = 1234,
    chains = 8,
    parallel_chains = 4,
    refresh = 500,
    iter_warmup = 1000, #3000,
    iter_sampling = 5000,
    thin =  1,
    adapt_delta = 0.99)
```

do some comparisons with parameter values
```{r}
estimated_locality_means = as.numeric(apply(fit$draws("mu_locality"),3,mean))

plot(c_l, estimated_locality_means);abline(0,1)
```
```{r}
estimated_reef_means = as.numeric(apply(fit$draws("mu_reef"),3,mean))
plot(nRL$reef_mean, estimated_reef_means); abline(0,1)
```
```{r}
print(c(estimated_sigma_locality = as.numeric(apply(fit$draws("sigma_locality"),3,median)),locality_sigma))
```

```{r}
print(c(estimated_sigma_reef = as.numeric(apply(fit$draws("sigma_reef"),3,median)),reef_sigma))
```
```{r}
print(c(estimated_sigma_hole = as.numeric(apply(fit$draws("sigma_hole"),3,median)),hole_sigma))
```

```{r}
print(c(estimated_depth = as.numeric(apply(fit$draws("d"),3,median)),depth_mean))
```
sigma_depth
```{r}
print(c(estimated_depth_sigma = as.numeric(apply(fit$draws("sigma_depth"),3,median)), depth_sigma))
```
mu_depth
```{r}
print(c(estimated_depth_mu = as.numeric(apply(fit$draws("mu_depth"),3,median)), depth_mean))
```

 
 
 

