---
title: "GeoChron"
output: html_notebook
---
Assess variability of age or CPE estimates at multiple levels.
At the highest level is a region which is composed of $nL$ localities.

The mean age across a region is $e$ which generates means $c_l$ for each locality.

$$ c_l = e + \zeta_c, l = 1,\ldots, nL, \zeta_c \sim N(0, \sigma_c^2)$$
Each locality has a number $nR_l$ of constituent reefs and a mean $c_l$ which generates means for each reef.

$$r_{kl} = c_l + \zeta_r, k = 1,\ldots, nR_l, l = 1\ldots nL, \zeta_r \sim N(0,\sigma_r^2)$$


Each reef has a number $nH_{kl}$ of constituent holes (typically 3 for our data) and the reef mean generates observations for each hole.

$$ x_{jkl} = r_{kl} + \zeta_h, j = 1,\ldots, nH_{kl}, k = 1,\ldots, nR_l, l = 1,\ldots, nL, \zeta_h \sim N(0,\sigma_h^2)$$
Finally, for each hole there are actuall two depths and one would expect the deeper sample to be older than the sample above it. We model the offset $d$ and report its posterior distribution.
We are interested in comparing its variance $\sigma_d^2$ with the other sigmas. 

$$ y_{jkl} = d + x_{jkl} $$
$$ y_{jkl}-x_{jkl} \sim N(\mu_d, \sigma_d^2)  $$

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(magrittr))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(cmdstanr))
suppressPackageStartupMessages(library(bayesplot))
```
 
Create data list from Geochronology "sampsum" data table created from "HOBS-geochronology-ms2.rmd". For a simulation illustrating how this model works, see "GeoChron_sim.Rmd."

```{r}
sampsum[, Loc := str_sub(sample, 1, 2), by = sample]
sampsum[, Locality := as.integer(factor(Loc, levels = unique(sampsum$Loc)))]
# sampsum[, reef2 := fcase(Locality %in% c(1, 2, 5:7, 9:11), reef,
#                          Locality == 3, ifelse(reef == 5, 4L, reef),
#                          Locality == 4, ifelse(reef == 4, 3L, reef),
#                          Locality == 8, as.integer(reef - 1)), by = c("Locality", "reef")]
sampsum[, reef3 := as.integer(factor(reefID, levels = unique(sampsum$reefID)))]
sampsum[, reef2Locality := Locality, by = c("Locality", "reef")]
sampsum[, holeID2 := holeID]
sampsum[, sampsperhole := length(holeID), by = holeID2]
sampsum_r2l <- distinct(sampsum[, .(reef, 
                                    locality, 
                                    reefID,
                                    Locality,
                                    reef3,
                                    reef2Locality)])

sampsum_15to25cm <- distinct(sampsum[depth == "15-25", .(sample,
                                                         Sample_median_age,
                                                         median_age_bp,
                                                         Sample_corrected_posterior_age_estimate,
                                                         reef,
                                                         locality,
                                                         reefID,
                                                         holeID,
                                                         holeID2,
                                                         Locality,
                                                         reef3,
                                                         reef2Locality)])

sampsum_25to35cm <- distinct(sampsum[depth == "25-35", .(sample,
                                                         Sample_median_age,
                                                         median_age_bp,
                                                         Sample_corrected_posterior_age_estimate,
                                                         reef,
                                                         locality,
                                                         reefID,
                                                         holeID,
                                                         holeID2,
                                                         Locality,
                                                         reef3,
                                                         reef2Locality)])

depthdiffs <- sampsum[sampsperhole == 2, .(sample, 
                                           Sample_median_age,
                                           median_age_bp,
                                           Sample_corrected_posterior_age_estimate, 
                                           depth, 
                                           reef, 
                                           locality, 
                                           reefID, 
                                           holeID,
                                           holeID2,
                                           Locality,
                                           reef3,
                                           reef2Locality)]

depthdiffs[, `:=` (depthdiff_mabp = median_age_bp[depth == "25-35"] -  median_age_bp[depth == "15-25"], 
                   depthdiff_cpe = Sample_corrected_posterior_age_estimate[depth == "25-35"] - Sample_corrected_posterior_age_estimate[depth == "15-25"]), by = holeID2]

depthdiffs2 <- distinct(depthdiffs[, .(reef, locality, reefID, holeID, Locality, reef3, reef2Locality, depthdiff_mabp, depthdiff_cpe)])

locs <- distinct(sampsum_15to25cm[, .(locality, Locality)])
locs[, loclab := 12 - Locality]
setorder(locs, "Locality")

reefs <- distinct(sampsum_15to25cm[, .(reefID, reef3)])
reefs[, reeflab := 32 - reef3]
setorder(reefs, "reef3")

#median estimated sample ages and corresponding time-averaging estimates from Dominguez et al. (2016)
Detal <- data.table(median_age_bp = c(123, 283, 146, 148, 160, 2043), 
                    Sample_corrected_posterior_age_estimate = c(68, 2670, 16, 27, 16, 1937),
                    Locality = c(1, 1, 1, 1, 1, 1),
                    Reef = c(1, 2, 3, 4, 5, 6),
                    Reef2Locality = c(1, 1, 1, 1, 1, 1),
                    nL = 1,
                    nR = 6)

#for median age over space (15-25cm burial depth)
HOBS_medagexspace_15to25cm <- list(N = nrow(sampsum_15to25cm),
                                   Locality = sampsum_15to25cm$Locality,
                                   Reef = sampsum_15to25cm$reef3,
                                   Reef2Locality = sampsum_r2l$reef2Locality,
                                   x = sampsum_15to25cm$median_age_bp,
                                   nL = length(unique(sampsum_15to25cm$Locality)),
                                   nR = length(unique(sampsum_15to25cm$reef3)))

#for CPE over space (15-25cm burial depth)
HOBS_cpexspace_15to25cm <- list(N = nrow(sampsum_15to25cm),
                                Locality = sampsum_15to25cm$Locality,
                                Reef = sampsum_15to25cm$reef3,
                                Reef2Locality = sampsum_r2l$reef2Locality,
                                x = sampsum_15to25cm$Sample_corrected_posterior_age_estimate,
                                nL = length(unique(sampsum_15to25cm$Locality)),
                                nR = length(unique(sampsum_15to25cm$reef3)))

#for median age over space (25-35cm burial depth)
HOBS_medagexspace_25to35cm <- list(N = nrow(sampsum_25to35cm),
                                   Locality = sampsum_25to35cm$Locality,
                                   Reef = sampsum_25to35cm$reef3,
                                   Reef2Locality = sampsum_r2l$reef2Locality,
                                   x = sampsum_25to35cm$median_age_bp,
                                   nL = length(unique(sampsum_25to35cm$Locality)),
                                   nR = length(unique(sampsum_25to35cm$reef3)))

#for CPE over space (25-35cm burial depth)
HOBS_cpexspace_25to35cm <- list(N = nrow(sampsum_25to35cm),
                                Locality = sampsum_25to35cm$Locality,
                                Reef = sampsum_25to35cm$reef3,
                                Reef2Locality = sampsum_r2l$reef2Locality,
                                x = sampsum_25to35cm$Sample_corrected_posterior_age_estimate,
                                nL = length(unique(sampsum_25to35cm$Locality)),
                                nR = length(unique(sampsum_25to35cm$reef3)))

#for median age depth differences over space
HOBS_medageddxspace <- list(N = nrow(depthdiffs2),
                            Locality = depthdiffs2$Locality,
                            Reef = depthdiffs2$reef3,
                            Reef2Locality = sampsum_r2l$reef2Locality,
                            x = depthdiffs2$depthdiff_mabp,
                            nL = length(unique(depthdiffs2$Locality)),
                            nR = length(unique(depthdiffs2$reef3)))

#for CPE depth differences over space
HOBS_cpeddxspace <- list(N = nrow(depthdiffs2),
                         Locality = depthdiffs2$Locality,
                         Reef = depthdiffs2$reef3,
                         Reef2Locality = sampsum_r2l$reef2Locality,
                         x = depthdiffs2$depthdiff_cpe,
                         nL = length(unique(depthdiffs2$Locality)),
                         nR = length(unique(depthdiffs2$reef3)))

#for median age depth differences over space - Dominguez et al. (2016) data
Detal_medagexspace <- list(N = 6,
                           Locality = Detal$Locality,
                           Reef = Detal$Reef,
                           Reef2Locality = Detal$Reef2Locality,
                           x = Detal$median_age_bp,
                           nL = Detal$nL,
                           nR = Detal$nR)

#for CPE depth differences over space - Dominguez et al. (2016) data
Detal_cpexspace <- list(N = 6,
                        Locality = Detal$Locality,
                        Reef = Detal$Reef,
                        Reef2Locality = Detal$Reef2Locality,
                        x = Detal$Sample_corrected_posterior_age_estimate,
                        nL = Detal$nL,
                        nR = Detal$nR)

```

```{r}

#For HOBS data
model_HOBS <- cmdstan_model("GeoChron.stan")

#For Dominguez et al. (2016) data
model_Detal <- cmdstan_model("GeoChron_Detal.stan")

```

```{r}

datalists <- c(as.name("HOBS_medagexspace_15to25cm"), 
               as.name("HOBS_cpexspace_15to25cm"),
               as.name("HOBS_medagexspace_25to35cm"),
               as.name("HOBS_cpexspace_25to35cm"),
               as.name("HOBS_medageddxspace"),
               as.name("HOBS_cpeddxspace"),
               as.name("Detal_medagexspace"),
               as.name("Detal_cpexspace"))

for(Data in datalists){
  if(str_detect(Data, "HOBS")){
    model <- cmdstan_model("GeoChron.stan")
  } else{
    model <- cmdstan_model("GeoChron_Detal.stan")
  }
  
  fit = model$sample(data = eval(Data),
        seed = 1234,
        chains = 8,
        parallel_chains = 8,
        refresh = 1000,
        iter_warmup = 3000, #3000,
        iter_sampling = 15000,
        thin =  3,
        adapt_delta = 0.99,
        max_treedepth = 15)
  
  fit$save_output_files("StanOutput")
  fit$cmdstan_diagnose()
  dim(fit$draws()) # accessing draws seems to fill the structure with draws
  saveRDS(fit, file = here::here(paste0("StanOutput/", Data, ".rds")))
  
  plot1 <- bayesplot::mcmc_intervals(fit$draws("mu_locality")) 
  plot1 <- plot1 +
    geom_text(data = locs, aes(x = min(plot1$data$ll) - 3, y = loclab, label = locality), inherit.aes = FALSE) +
    xlim(min(plot1$data$ll) - 5, max(plot1$data$hh))
  
  saveRDS(plot1, here::here(paste0("StanOutput/", Data, "locintplot.rds")))
  ggsave(filename = here::here(paste0("StanOutput/", Data, "locintplot.pdf")),
         plot = plot1,
         width = 6,
         height = 6,
         units = "in",
         dpi = 400)
  
  plot2 <- bayesplot::mcmc_intervals(fit$draws("mu_reef"))
  plot2 <- plot2 +
    geom_text(data = reefs, aes(x = min(plot2$data$ll) - 10, y = reeflab, label = reefID), inherit.aes = FALSE) +
    xlim(min(plot2$data$ll) - 12, max(plot2$data$hh))
  
  saveRDS(plot2, here::here(paste0("StanOutput/", Data, "reefintplot.rds")))
  ggsave(filename = here::here(paste0("StanOutput/", Data, "reefintplot.pdf")),
         plot = plot2,
         width = 6,
         height = 6,
         units = "in",
         dpi = 400)
}

```


```{r}

bayesplot::mcmc_dens(fit$draws("mu_locality"), facet_args = list(scales = "fixed", ncol = 3))
plot1 <- bayesplot::mcmc_intervals(fit$draws("mu_locality")) 
plot1 +
  geom_text(data = locs, aes(x = min(plot1$data$ll) - 3, y = loclab, label = locality), inherit.aes = FALSE) +
  xlim(min(plot1$data$ll) - 5, max(plot1$data$hh))
  

```

```{r}
plot2 <- bayesplot::mcmc_intervals(fit$draws("mu_reef"))
plot2 +
  geom_text(data = reefs, aes(x = min(plot2$data$ll) - 8, y = reeflab, label = reefID), inherit.aes = FALSE) +
  xlim(min(plot2$data$ll) - 10, max(plot2$data$hh))
```

 
 
 

