---
title: "New Pellicer Creek"
output: html_notebook
---

DEP AGREEMENT NO. CZ325 Deliverable 2b (Radiocarbon sample size for age and time-averaging determination)
```{r}
library(tidyverse)
library(magrittr)
library(reshape2)
library(HDInterval)
library(readxl)
library(oxcAAR)
setOxcalExecutablePath("/Users/johnhandley/Documents/PaleontologyResearch/Oysters/HOBS/Contract work/OxCal/bin/OxCalMac")
```

"Utilities.R" is part ofDEP AGREEMENT NO. CZ325, Deliverable 1a (Modeling approach to age and time-averaging estimation)
```{r}
source("Utilities.R") 
```

Helper function to create calendar dates from fractions modern
```{r}
source("write_oxcal_sum_script.R")
source("SetUpNewCurve.R") # sets up globals curvename, curverpath

oxcARR_Calibration = function(fractions, sample_name = "TEST", curvename = "HOBS_14C_CalCurve_PredictedVals_20230524",curvepath = "/Users/johnhandley/Documents/PaleontologyResearch/Oysters/HOBS/Contract work/Comparisons/Kowalewski_NP/Curves/HOBS_14C_CalCurve_PredictedVals_20230524.14c"){

shell_names = unique(fractions$name)
code = write_oxcal_sum_script(fractions, path = "OxCal_Sum_scripts", filename = sample_name,curvename = curvename,curvepath = curvepath)
    
    
    my_result_file <- executeOxcalScript(code) # looks like it stores it in JSON
    my_result_text <- readOxcalOutput(my_result_file)
    
    #Or you get the whole output of Oxcal as object:
    
    #my_result_data <- parseFullOxcalOutput(my_result_text)
    my_result_data = parseOxcalOutput(my_result_text, only.R_Date = F)
    
    local_posterior_distribution = data.frame() # one per object
    for( p in my_result_data) {
      
        if( p$name %in% c(shell_names,"Sum"))
        {
          df = data.frame(year = p$raw_probabilities$dates,
                          probability = p$raw_probabilities$probabilities/sum( p$raw_probabilities$probabilities))
          df %<>% mutate(name = p$name)
          local_posterior_distribution = rbind(local_posterior_distribution, df)
        }
      
    }
    
return(local_posterior_distribution)
}
```

Get the new samples

```{r}
Durham_GAS_April24_2023 = read_xlsx("Durham_GAS 14C results_April 24 2023.xlsx", skip = 1)
Durham_GAS_April24_2023 %<>% select("Sample ID", "Modern","±...5")
Durham_GAS_April24_2023 %<>% na.omit()
names(Durham_GAS_April24_2023) = c("name","f14C", "f14C_sd")
Durham_GAS_April24_2023 %<>% mutate(Sample_code = as.character(data.frame(strsplit(name, " "))[1,]))
```

```{r}
SpecNames_Xwalk = read.csv("SpecNames_Xwalk.csv")

```

From the new Pellicer specimens, find the old ones from the sample location (Reef, Sample), e.g., PC_R3H2S2 26
```{r}
Durham_GAS_April24_2023 %>% filter(grepl("PC_R3H2S2", name)) -> PC_New

```


Looks like Station 8933
```{r}
SpecNames_Xwalk %>% filter(grepl("PC_R3H2S2", Sample_ID))
```

Get the original HOBS data; S2 corresponds to "25-35cm"
```{r}
HOBSGeo_Apps_202208_final = read_xlsx("HOBSGeo_Apps_202208_final.xlsx", sheet = "DR3", skip = 31 )
HOBSGeo_Apps_202208_final %>% filter(Locality == "Pellicer Creek" & Depth == "25-35cm" ) -> PC_Old
PC_Old %<>% filter(Station == 8933) %>% select(Cat_No,F14C_corr,F14C_corr_sd,Depth,Station)
names(PC_Old) = c("name", "F14C_corr","F14C_corr_sd","Depth","Station")
```

Run the old through the Calibration
```{r}
PC_Old_posteriors = oxcARR_Calibration(PC_Old, sample_name = "PC_Old")
```

```{r}
PC_Old_posteriors %>% filter(name != "Sum") %>% ggplot(aes(x = year, y = probability, color = name)) + geom_line(show.legend = FALSE) + xlim(c(1100,2020))
```

Run the new through the Calibration
```{r}
PC_New_posteriors = oxcARR_Calibration(PC_New, sample_name = "PC_New")
```

```{r}
PC_New_posteriors %>% filter(name != "Sum") %>% ggplot(aes(x = year, y = probability, color = name)) + geom_line(show.legend = FALSE)+ xlim(c(1100,2020))
```

```{r}
Pellicer_Creek_R3H2S2 = rbind(PC_Old_posteriors %>% filter(name != "Sum"),PC_New_posteriors %>% filter(name != "Sum"))
```

```{r}
 Pellicer_Creek_R3H2S2_names = unique((Pellicer_Creek_R3H2S2  %>% select(name))$name) 
 Pellicer_Creek_R3H2S2_results = data.frame()
 set.seed(1234)
for( n in 2:length(Pellicer_Creek_R3H2S2_names)){
     for( ii in 1:100){
        random_sample = sample(Pellicer_Creek_R3H2S2_names, size = n, replace = FALSE)
        tmp = Pellicer_Creek_R3H2S2 %>% filter(name %in% random_sample)
        Pellicer_Creek_R3H2S2_results = rbind(Pellicer_Creek_R3H2S2_results
            , data.frame( n = n, cpe = CPE(tmp)$cpe
                          , tav_iqr =CPE(tmp)$tav_iqr
                          , cr_0.95 = CR_95_TAV(tmp)
                          , tav_hdi = HDI_95_TAV(tmp)
                          , hdi_of_hdi = HDI_of_HDI(tmp, credMass = 0.5)
                          , Median = Median(TAV(tmp)) )) #, ETA = ETA(tmp) ))
     }
}
```

```{r} 
 write_rds(Pellicer_Creek_R3H2S2_results,"Pellicer_Creek_R3H2S2_results.RDS")
```

Original Shells, new calibration curve
```{r}
Old_posteriors = PC_Old_posteriors
Old_fractions = PC_Old
PC_Old_results = data.frame(n = nrow(Old_fractions), cpe = CPE(Old_posteriors)$cpe
                            , tav_iqr =  CPE(Old_posteriors)$tav_iqr
                            , Median = Median(TAV(Old_posteriors))
                            , cr_0.95 = CR_95_TAV(Old_posteriors)
                            , tav_hdi = HDI_95_TAV(Old_posteriors)
                            , hdi_of_hdi = HDI_of_HDI(Old_posteriors, credMass = 0.5))
```
This is the results for the original data but new curve
```{r}
write_rds(PC_Old_results, "Pellicer_Creek_Old_Data_New_Curve_results.RDS")
```

Original results with original curve
```{r}
HOBSGeo_Apps_202208_final.DR1 = read_xlsx("HOBSGeo_Apps_202208_final.xlsx", sheet = "DR1", skip = 21 )
names(HOBSGeo_Apps_202208_final.DR1) = c("Locality",	"Station",	"Lat",	"Lon",	"IGSN",	"Depth",	"N", "Median",	"Min",	"Max","IQR_AER",	"IQR_TAV",	"IQR_CPE")
HOBSGeo_Apps_202208_final.DR1 %>% filter(Station == 8933 & Depth == "25-35cm") %>% select(N, Median,IQR_TAV,IQR_CPE) %>% data.frame() -> PC_original_estimates
```

Original Data, Original Curve
```{r}
write_rds(PC_original_estimates, "PC_original_estimates.RDS")
```


```{r}
Pellicer_Creek_R3H2S2_results %>% select(-Median) %>% reshape2::melt("n") %>% ggplot(aes(x = n, y = value)) + geom_jitter(width = 0.2) + geom_smooth(color = "orange") + facet_wrap(~variable) + ggtitle("Pellicer Creek")
```
```{r}
Pellicer_Creek_R3H2S2_results %>% select(n,Median) %>% reshape2::melt("n") %>% ggplot(aes(x = n, y = value)) + geom_jitter(width = 0.2, alpha =0.1, size = 0.5) + geom_smooth(color = "orange") + geom_point(data = PC_original_estimates, aes(x = N, y = Median,color = "Original Curve"), size = 2) + geom_point(data = PC_Old_results, aes(x = n, y = Median, color = "New Curve"), size = 2) + ggtitle("Pellicer Creek") + ylab("Median")
```

```{r}
Pellicer_Creek_R3H2S2_results %>% select(n,cpe) %>% reshape2::melt("n") %>% ggplot(aes(x = n, y = value)) + geom_jitter(width = 0.2, alpha =0.1, size = 0.5) + geom_smooth(color = "orange") + geom_point(data = PC_original_estimates, aes(x = N, y = IQR_CPE,color = "Original Curve"), size = 2)  + geom_point(data = PC_Old_results, aes(x = n, y = cpe, color = "New Curve"), size = 2) + ggtitle("Pellicer Creek") + ylab("IQR_CPE")
```

```{r}
Pellicer_Creek_R3H2S2_results %>% select(n,tav_iqr) %>% reshape2::melt("n") %>% ggplot(aes(x = n, y = value)) + geom_jitter(width = 0.2, alpha =0.1, size = 0.5) + geom_smooth(color = "orange") + geom_point(data = PC_original_estimates, aes(x = N, y = IQR_TAV,color = "Original Curve"), size = 2) + geom_point(data = PC_Old_results, aes(x = n, y = tav_iqr, color = "New Curve"), size = 2)  + ggtitle("Pellicer Creek") + ylab("IQR_TAV")
```

```{r}
Pellicer_Creek_R3H2S2_results %>% select(n,cr_0.95) %>% reshape2::melt("n") %>% ggplot(aes(x = n, y = value)) + geom_jitter(width = 0.2, alpha =0.1, size = 0.5) + geom_smooth(color = "orange")  + geom_point(data = PC_Old_results, aes(x = n, y = cr_0.95, color = "New Curve"), size = 2)  + ggtitle("Pellicer Creek") + ylab("cr_0.95")
```

```{r}
Pellicer_Creek_R3H2S2_results %>% select(n,tav_hdi) %>% reshape2::melt("n") %>% ggplot(aes(x = n, y = value)) + geom_jitter(width = 0.2, alpha =0.1, size = 0.5) + geom_smooth(color = "orange")  + geom_point(data = PC_Old_results, aes(x = n, y = tav_hdi, color = "New Curve"), size = 2)  + ggtitle("Pellicer Creek") + ylab("tav_hdi")
```

```{r}
Pellicer_Creek_R3H2S2_results %>% select(n,hdi_of_hdi) %>% reshape2::melt("n") %>% ggplot(aes(x = n, y = value)) + geom_jitter(width = 0.2, alpha =0.1, size = 0.5) + geom_smooth(color = "orange")  + geom_point(data = PC_Old_results, aes(x = n, y = hdi_of_hdi, color = "New Curve"),  size = 2)  + ggtitle("Pellicer Creek") + ylab("hdi_of_hdi")
```


```{r}
pop_values = last(Pellicer_Creek_R3H2S2_results)
cpe.tmp = Pellicer_Creek_R3H2S2_results %>% group_by(n) %>% dplyr::summarize( cpe_Bias = Bias(cpe, pop_values$cpe), cpe_SD = StandardDeviation(cpe, pop_values$cpe), cpe_CV =  StandardDeviation(cpe, pop_values$cpe)/pop_values$cpe)

tav_iqr.tmp = Pellicer_Creek_R3H2S2_results %>% group_by(n) %>% dplyr::summarize( tav_iqr_Bias = Bias(tav_iqr, pop_values$tav_iqr), tav_iqr_SD = StandardDeviation(tav_iqr, pop_values$tav_iqr), tav_iqr_CV =  StandardDeviation(tav_iqr, pop_values$tav_iqr)/pop_values$tav_iqr) %>% select(-n)

cr_0.95.tmp = Pellicer_Creek_R3H2S2_results %>% group_by(n) %>% dplyr::summarize( cr_0.95_Bias = Bias(cr_0.95, pop_values$cr_0.95), cr_0.95_SD = StandardDeviation(cr_0.95, pop_values$cr_0.95), cr_0.95_CV =  StandardDeviation(cr_0.95, pop_values$cr_0.95)/pop_values$cr_0.95)%>% select(-n)

tav_hdi.tmp = Pellicer_Creek_R3H2S2_results %>% group_by(n) %>% dplyr::summarize( tav_hdi_Bias = Bias(tav_hdi, pop_values$tav_hdi), tav_hdi_SD = StandardDeviation(tav_hdi, pop_values$tav_hdi), tav_hdi_CV =  StandardDeviation(tav_hdi, pop_values$tav_hdi)/pop_values$tav_hdi)%>% select(-n)

hdi_of_hdi.tmp = Pellicer_Creek_R3H2S2_results %>% group_by(n) %>% dplyr::summarize( hdi_of_hdi_Bias = Bias(hdi_of_hdi, pop_values$hdi_of_hdi), hdi_of_hdi_SD = StandardDeviation(hdi_of_hdi, pop_values$hdi_of_hdi), hdi_of_hdi_CV =  StandardDeviation(hdi_of_hdi, pop_values$hdi_of_hdi)/pop_values$hdi_of_hdi)%>% select(-n)

Pellicer_Creek_R3H2S2_summary = round(cbind(cpe.tmp,tav_iqr.tmp,cr_0.95.tmp,tav_hdi.tmp,hdi_of_hdi.tmp), 3)
Pellicer_Creek_R3H2S2_summary
```



```{r}
Pellicer_Creek_R3H2S2_summary %>% select(n,contains("Bias")) %>% melt("n") %>% ggplot(aes(x = n, y = value, color = variable)) + geom_line()+ ylab("Bias") + ggtitle("Pelicer Creek")
```
```{r}
Pellicer_Creek_R3H2S2_summary %>% select(n,contains("SD")) %>% melt("n") %>% ggplot(aes(x = n, y = value, color = variable)) + geom_line() + ylab("SD")+ ggtitle("Pelicer Creek")
```

```{r}
Pellicer_Creek_R3H2S2_summary %>% select(n,contains("CV")) %>% melt("n") %>% ggplot(aes(x = n, y = value, color = variable)) + geom_line()+ ylab("CV")+ ggtitle("Pelicer Creek")
```
